Molecular dynamics simulation of synchronization in driven particles

### 2021-02-09 manuscript/manuscript_ajp_2020.tex
moved phase plots description from introduction to exercises

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
with 
Huygens' 1665 experiments on
motions of 
pendula in two wall-mounted clocks.
Even when started out of phase, 
%Huygens demonstrated the tendency of
coupled pendula tend to swing in time
due to interactions through the wall \cite{Bennett2002}.
%phase locking
A locked-mode is an integer frequency ratio.
In Huygens clocks, 
the pendula were observed swing at same rate in a 1:1 mode.
For simple pendula of different lengths,
a 2:1 mode can be seen when a pendulum four
times the length of the other swings with twice the period.
%
%to a mode or frequency ratio  
%of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena
Synchronization 
%is often studied with simple computational models and
can be visualized 
with phase
plots or Lissajous figures, 
first reported in 1857 \cite{Lissajous1857}.
%and easily generated with a computer code 
%or oscilloscope \cite{Tong1997}. 
A mode-locked system is confined to 
a closed loop in  
parametric phase space % showing the relationship between two periodic functions
with patterns determined by the mode. % frequency ratio,
({\it As we will demonstrate in phase plots in Figs. X, Y, Z}).
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
Note the term phase
has two scientific meanings:  
%``oscillation phase'' and ``phase space.''
%refer to different uses of the word ``phase.''
oscillation phase $\phi_0$ is the constant of the argument
in a periodic function
$\sin{(\omega t + \phi_0)}$ and phase space is
used to characterize oscillating systems with
parametric plots of $v_y$ vs. $y$.

@@

Light is a
tool for manipulating the colloidal environment.
A circular optical trap uses 
radiation pressure from 
a laser beam to confine colloids \cite{Ashkin1997}.
%Since photon density is high at the beam center,
A colloid centered in the beam is %subject to a
uniformly bombarded by photons while 
%form a circular pattern with a higher density 
%central 
%energy minimum.
%optical trap  trap colloids
%with radiation pressure  which uses  photon 
%control colloid motions 
%A  is designed
off-center colloids %away from center
collide with photons unevenly,
creating a net force.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either moves colloids toward the local minimum 
or ejects it from the trap.
An optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces,
and has been studied in detail in Ref.~\cite{Volpe2013}.
%In experiments
Diffraction gratings can be used to create
more complex light environments, 
%with one or more laser beams,
such as periodic patterns of minima
suitable for synchronization studies \cite{Grier2003}.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
%In Section~\ref{sec:colloids} we describe colloid particles,
%an relevant
%experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
%In Section~\ref{sec:sync}
%we 
%extend the model to include multiple particles,
%and show how confined 
%interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
We include 
exercises for interested students 
in Section~\ref{sec:problems}.
In Section~\ref{sec:conclusion}
we describe how our work with colloids applies more broadly
to many physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.

@@

  \subsection{Visualizing motion + phase plots}
  \label{ex:animation}

@@

  Phase plots...

%\end{enumerate}

>>>>

%Christiaan Huygens 
%[has been adapted into many useful technologies...]
Synchronized phase-locking or mode-locking 
first appeared in the scientific literature
with 
Huygens' 1665 experiments on
motions of synchronized 
pendula in wall-mounted clocks \cite{Bennett2002}.
%Even when started out of phase, 
%Huygens demonstrated the tendency of
%coupled pendula tend to swing in time
%due to interactions through the wall 
%phase locking
A locked-mode is an integer frequency ratio.
In Huygens clocks,
%even when started out of phase, 
the pendula were observed swing at same rate -- a 1:1 mode.
For simple pendula of different lengths,
a 2:1 mode can be seen when a pendulum four
times the length of another swings with twice the period.
%
%to a mode or frequency ratio  
%of an integer number \cite{Bak1986}.
%Dynamical mode locking
%The phenomena

@@

Light is a
tool for manipulating the colloidal environment.
A circular optical trap uses 
radiation pressure from 
a laser beam to confine colloids \cite{Ashkin1997}.
%Since photon density is high at the beam center,
A colloid centered in the beam is %subject to a
uniformly bombarded by photons while 
%form a circular pattern with a higher density 
%central 
%energy minimum.
%optical trap  trap colloids
%with radiation pressure  which uses  photon 
%control colloid motions 
%A  is designed
off-center colloids %away from center
collide with photons unevenly,
creating a net force.
%Colloids at a trap edge
%experience a net force
%due to the 
Depending on the %shape of the trap and
location of the particle in the trap,
the radiation pressure either moves colloids toward the beam center 
%local minimum 
or ejects it from the trap.
An optically trapped colloid executing Brownian motion
is a useful probe of microscopic forces ~\cite{Volpe2013}.
%In experiments
Diffraction gratings can be used to create
more complex light environments, 
%with one or more laser beams,
such as periodic patterns of minima
suitable for synchronization studies \cite{Grier2003}.

@@

%Given the prevalence of synchronization,
%it is the subject of many computational models.
Here we perform numerical studies 
on the synchronized dynamics
of confined particles driven over
a washboard shaped potential energy landscape.
%Numerical studies %of harmonic systems
%can yield insights
%to the mechanics of naturally oscillating systems.
We chose this model for its
relevance to condensed matter systems
and ease of simulation.
%---------------------------------------------------------------
%kinks and nanofriction
%-----------------------------------------------------------------
%transition from synchronized to chaotic
%describe types of physical systems - particle types
%%In this work,
%we perform 
%numerical simulations of confined, driven particles
%to model a variety of physical phenomena.
%-----------------------------------------------------------------------
%paper outline
%In the following paper
%In Section~\ref{sec:colloids} we describe colloid particles,
%an relevant
%experimental analog of our simulations.
We describe
our molecular dynamics model for a single particle in Section~\ref{sec:MD}.
%In this section, we restrict our discussion to a single particle,
%and include additional information
%for multiple particles in Section~\ref{sec:sync}.
%including various  and confining environments.
We summarize
our results 
%in Section~\ref{sec:results} 
including synchronized motion of a single confined particle
driven across a periodic landscape in 
Section~\ref{sec:one}.
%In Section~\ref{sec:sync}
%we 
%extend the model to include multiple particles,
%and show how confined 
%interacting particles respond to the same applied driving force.
%including stationary propagation of high density kinks
%in Sec.~\ref{sec:kink}.
%We include code to simulate
%and visualize the landscape and dynamics in this section
%and supplementary material.
%we
%demonstrate how uniform environments and applied forces
%and create synchronized flow patterns.
%We present these results using standard tools of non-linear oscillators
%such phase diagrams of velocity versus position.
%PHASE DIAGRAMS SHOWING COUPLING, OR QUASI-PERIODIC COUPLING
%INDICATE COOPERATIVE MOTION IN THE Q1D SUBSTRATE
%In Section~\ref{sec:quasiperiod},
%we show how an aperiodic landscape modifies the particle dynamics.
%Finally we explore the transition to chaotic dynamics in 
%Section~\ref{sec:chaos}
%and conclude 
We include 
exercises for interested students 
in Section~\ref{sec:problems}.
In Section~\ref{sec:conclusion}
we describe how our results apply 
to physical systems
such as dusty plasmas, superconducting vortices and Josephson junctions.

@@

\subsection{Phase plots}
  \label{ex:phase}

  Synchronization 
  %is often studied with simple computational models and
can be visualized 
with phase
plots or Lissajous figures, 
first reported in 1857 \cite{Lissajous1857}.
%and easily generated with a computer code 
%or oscilloscope \cite{Tong1997}. 
A mode-locked system is confined to 
a closed loop in  
parametric phase space % showing the relationship between two periodic functions
with patterns determined by the mode. % frequency ratio,
({\it As we will demonstrate in phase plots in Figs. X, Y, Z}).
%with a frequency generator,
%where two combined frequencies at perpendicular orientations
Note the term phase
has two scientific meanings:  
%``oscillation phase'' and ``phase space.''
%refer to different uses of the word ``phase.''
oscillation phase $\phi_0$ is the constant of the argument
in a periodic function
$\sin{(\omega t + \phi_0)}$ and phase space is
used to characterize oscillating systems with
parametric plots of $v_y$ vs. $y$.

  \subsection{Visualizing motion}
  \label{ex:animation}

@@
